Extracting Application and Menu Events from UI Tutorial Videos

### 2015-05-11 paper/main.tex
Cleaned up first two sections, also deleted some info from references to cut down bibliography size

In this project, we present two main contributions. First, we show that a standard neural network
following the AlexNet architecture with minimal tweaking can accurately recognize frames belonging
to a class of videos.  Second, we present a pipeline and procedure for detecting menus within a
video frame sequence. We then discuss how these two stages might be combined to enhance the process
of extracting useful information from videos.

@@

The proliferation of video tutorials on using different software raises new research questions on
how to improve their use to help users learn software~\cite{matejka_ambient_2011,
pongnumkul_pause-and-play_2011}.  One strength of video tutorials may be that users can directly
observe expert use, which can serve as a strong learning aid. Grossman \& Fitzmaurice showed
benefits of brief contextual video clips: short, 10-25 second contextually available videos were
effective to help users accomplish task and retain what they learned than traditional text-based
tutorials~\cite{grossman_toolclips_2010}. Unfortunately, video tutorials have drawbacks. Navigation
issues could lead to misunderstanding of content, and users may not be able to keep up with the pace
of instruction. Researchers working with longer, 2-3 minute task-oriented tutorials found that users
performed better with text tutorials than videos because users could not work at the same pace as
the video~\cite{grabler_generating_2009}.

In general, research on instructional videos point shows a need to segmented videos to emphasize the
steps of the task. One way to do that might be to develop techniques for detecting locations of
certain menu actions in the video, such as when the video shows its user clicking on a button in a
Photoshop video to create a blank image. Our goal, therefore, is to utilize computer vision
techniques to aid in this process.

@@

By examining pixels, many have succeeded at reverse engineering user interfaces and interactions
with user interfaces with purely visual information, independent of any knowledge of underlying
application implementation. Work by Zettleymoyer et
al.~\cite{zettlemoyer_ibots_1999}~\cite{zettlemoyer_visual_1999} examine widget identification in
IBOTS and VisMap.  Also
see~\cite{potter_triggers_1992}~\cite{amant_image_2005}~\cite{lieberman_visual_2001}.  Visual
identification of widgets and UIs may also have also been used in Olsen et al.'s ScreenCrayons work
that links annotations to arbitrary screen elements~\cite{olsen_jr_implementing_1999}, and in Tan et
al.'s work to interactively subdivide windows with a copy-paste metaphor~\cite{tan_wincuts_2004}.
\textbf{Andrew: Check on this.}

More recently, Sikuli~\cite{yeh_sikuli_2009} uses computer vision to identify GUI elements from
screen captures, using template matching for small UI elements and voting based on invariant local
features for large elements.  In application, Sikuli allows users to search for documentation
related to screenshots taken of UI elements, and to script GUI tasks using visual cues of the UI.
For searching based on screenshots, 3 types of features are extracted: text from the source
document, visual words computed from visual words from MSER-detected salient elliptical patches, and
3-grams over widget-embedded text extracted through OCR.  For searching on the screen for saved
screenshot images, small patterns are found via template matching and large patterns through
invariant feature voting.

Pause-and-Play extends the template matching approach of Sikuli to detect tool-selection UI events
in compressed, low-resolution video tutorials.  The system takes a tool palette template and tool
icons as input, producing metadata file of tool changes with timecodes.  With template matching at
multiple resolutions, detection is invariant to recording resolution and some camera effects.  Their
system cannot be run in in real-time~\cite{pongnumkul_pause-and-play_2011}.

A number of different technologies attempt to learn templates for arbitrary UI
elements~\cite{chang_associating_2011}~\cite{dixon_prefab_2010}~\cite{hurst_automatically_2010}~\cite{matejka_ambient_2011}.
Learning may require access to mouse APIs~\cite{hurst_automatically_2010} and accessibility
APIs~\cite{chang_associating_2011}.

Prefab identifies GUI elements with GUI specific visual features~\cite{dixon_prefab_2010}, which
enables overlay of advanced interactions on existing
interfaces~\cite{dixon_content_2011}~\cite{dixon_general-purpose_2012}.  \textbf{Andrew: Check up on
this initial description.  I'm also not sure if Prefab actually belongs in this conceptual group.}
To delve into the technical implementations of the Prefab pixel-based matching algorithm, we report
from~\cite{dixon_general-purpose_2012}.  Prefab identifies interface elements using a library of
prototypes.  This uses two strategies: exact matches of prototype pixels against an image, and
modelining the prototype background and differencing pixel in an image to identify foreground
interface elements.  Prefab generalizes from example images to models which contain parts like exact
pixel features, or regions or areas of variable size or repeating patterns.  It organizes the entire
interface into a hierarchy.

@@

\section{Video Classification}

This will be Daniel's section. I'm probably not going to include any subsections to save time.

Figure~\ref{fig:alex_net} shows our accuracy results for the two AlexNet networks we trained.

Figure~\ref{fig:per_class} shows our per-class accuracy results for the ``no noise'' AlexNet
network.

@@

Figures~\ref{fig:filters_all} and~\ref{fig:filters_nonoise} show visualizations of the 96 filters in
the first layer of AlexNet, each of which has dimension $55\times 55$. These were taken for the
model that had 1000 iterations of training.

@@

\section{Locating Menus in a Frame Sequence}

@@

\subsection{Data}

@@

\subsection{Implementation}
Here's a figure of our (edit: Andrew's) pipeline.  \textbf{Andrew: Add figure of pipeline.}

\subsection{Menu Location Results}

>>>>

In this project, we present two main contributions. First, we show that a standard neural network
following the AlexNet architecture with minimal tweaking can accurately recognize frames belonging
to a class of videos.  Second, we present a pipeline for detecting menus within a video frame
sequence. We then discuss how these two stages might be combined to enhance the process of
extracting information from videos.

@@

The proliferation of video tutorials on using different software raises new research questions on
how to improve users' ability to learn from them~\cite{matejka_ambient_2011,
pongnumkul_pause-and-play_2011}.  One strength of video tutorials is that users can directly observe
expert use, which can serve as a strong learning aid. Grossman \& Fitzmaurice showed that short,
10-25 second contextually-related videos were more effective to help users accomplish tasks than
traditional text-based tutorials~\cite{grossman_toolclips_2010}.  Unfortunately, video tutorials
have drawbacks. Navigation issues could lead to misunderstanding of content, and users may not be
able to keep up with the pace of instruction. Researchers working with longer, 2-3 minute
task-oriented tutorials found that users performed better with text tutorials than videos because
users could not work at the same pace as the video~\cite{grabler_generating_2009}.

In general, research on instructional videos point shows a need to segment videos to emphasize
certain steps. One way to automatically segment existing videos is to develop techniques for
detecting locations of certain menu actions in videos. Our goal, therefore, is to do this by
utilizing computer vision techniques.

@@

By examining pixels, many have succeeded at reverse engineering user interfaces and interactions
with purely visual information, independent of any knowledge of underlying application
implementations. For instance, Sikuli~\cite{yeh_sikuli_2009} uses computer vision to identify GUI
elements from screen captures, using template matching for small UI elements and voting based on
invariant local features for large elements. Sikuli allows users to search for documentation related
to screenshots of UI elements, and to script GUI tasks using visual cues.

Pause-and-Play extends the template matching approach of Sikuli to detect tool-selection UI events
in compressed, low-resolution video tutorials.  The system takes a tool palette template and tool
icons as input, producing metadata file of tool changes with timecodes.  With template matching at
multiple resolutions, detection is invariant to recording resolution and some camera
effects~\cite{pongnumkul_pause-and-play_2011}.

% Can cite these but we need more space for actual content: ~\cite{dixon_content_2011}~\cite{dixon_general-purpose_2012}.
Prefab identifies GUI elements with GUI specific visual features~\cite{dixon_prefab_2010}, which
enables overlay of advanced interactions on existing interfaces.  Prefab identifies interface
elements using a library of prototypes.  This uses two strategies: exact matches of prototype pixels
against an image, and modelining the prototype background and differencing pixel in an image to
identify foreground interface elements.  Prefab generalizes from example images to models which
contain parts like exact pixel features, or regions or areas of variable size or repeating patterns.
It organizes the entire interface into a hierarchy.

@@

\section{Video Classification}\label{sec:daniel}

As a first step in the information extraction pipeline, we consider the following classification
question: given a video tutorial, what application is it about? As an extra challenge, we will only
assume we have access to the raw frames, and not other meta data such as the video title, etc.
(Knowing that a video is called ``Eclipse Tutorial \#5'' would make the task too easy.) The
rationale for analyzing this classification problem is that if we know the application that forms
the basis of the video tutorial, then we can further specialize information extraction procedures.
For instance, we might develop templates for matching Eclipse menu items, which would be deployed
only if we are reasonably confident that a video tutorial is actually about Eclipse. Furthermore, if
a video contains multiple applications (e.g., a video might contain Excel, Powerpoint, and other
Microsoft Office products), we might be able to segment the video based on application type, which
relates to the need to partition videos.

@@

To collect training data, we used the open source \texttt{youtube-dl} software to download videos
from YouTube into mp4 format. We restricted our focus to videos about Eclipse, Excel, Photoshop, and
SketchUp, and we extracted from playlists to maintain consistency among the four classes. Most
videos were four to fifteen minutes long.  Then, we used the open source \texttt{ffmpeg} software to
extract one frame per second from each video, which resulted in a set of about 110240 JPG images,
roughly evenly split among the four classes. We randomly divided this into training, validation, and
testing sets, with 52000 training images (13000 per class), 8000 validation images (2000 per class),
and 1000 testing images (250 per class). All images were resized to $256\times 256$ pixels.

We observed that many of the videos we found had specialized beginning and ending scenes that
created frames that were different from what we would expect for those videos. For instance, the
SketchUp videos we found had a lengthy 10-15 second introductory scene that was the same for all
videos, but was not very similar to the various scenes inside the actual tutorial portion of the
videos. Thus, we created two datasets: one with all the frames we extracted, and one with the
\emph{first and last ten frames removed} for each video. Since we had one frame per second, this was
equivalent to ``cropping'' each video by ten seconds on both sides. This cropped dataset, which also
included cropped validation and testing frames, was termed the ``no noise'' data.

To train our network, we used CAFFE's pre-provided AlexNet architecture, but increased the momentum
from 0.9 to 0.99 and decreased the base learning rate from 0.01 to 0.001. We had fairly small batch
size of 32 images, so our updates are relatively noisy. We ran training for 1000 iterations on both
datasets and saved every 20th architecture for testing purposes.

Figure~\ref{fig:alex_net} shows our accuracy results for the two AlexNet networks we trained on the
test set of 1000 images. One can see that the architecture trained on all the data performed
inconsistently on its testing data, and its accuracy eventually leveled off at 25 percent since it
always guessed that frames belonged to SketchUp videos. In contrast, the architecture trained and
tested on ``noiseless'' data performed about as good as we could have hoped for, with accuracy
peaking at 96.1 percent after 1000 iterations.

We also decided to investigate per-class accuracy results to see if certain types of video frames
were harder to learn than others. Figure~\ref{fig:per_class} shows our per-class accuracy results
for the ``no noise'' AlexNet network. For the first few iterations, accuracy varies wildly among the
classes since it is not uncommon for the network to always pick one class. After that, it seems
clear that the SketchUp videos are the hardest to learn since the black curve regularly shows worse
accuracy results than the others. Fortunately, by the time we reach about 500 iterations, our
network can correctly classify the vast majority of images across all four classes.

We then investigated the 96 filters for the first layer of the AlexNet architecture, each of which
has dimension $55\times 55$. Figures~\ref{fig:filters_all} and~\ref{fig:filters_nonoise} show
visualizations for the two architectures taken after 1000 iterations of training. There are some
interesting patterns present, such as straight lines and rough, curved shapes.

@@

\section{Locating Menus in a Frame Sequence}\label{andrew}

@@

@@

Here's a figure of our (edit: Andrew's) pipeline.  \textbf{Andrew: Add figure of pipeline.}
